Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation with Real-Time Maude

### 2016-10-25 RMS/final-version/main.tex
initialization of final version

\documentclass[10pt,journal]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[10pt,journal,compsoc]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  \usepackage[nocompress,noadjust]{cite}
\else
  % normal IEEE
  \usepackage[noadjust]{cite}
\fi
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.
%
% Note that some packages require special options to format as the Computer
% Society requires. In particular, Computer Society  papers do not use
% compressed citation ranges as is done in typical IEEE papers
% (e.g., [1]-[4]). Instead, they list every citation separately in order
% (e.g., [1], [2], [3], [4]). To get the latter we need to load the cite
% package with the nocompress option which is supported by cite.sty v4.0
% and later. Note also the use of a CLASSOPTION conditional provided by
% IEEEtran.cls V1.7 and later.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/tex-archive/info/epslatex/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

\usepackage{algpseudocode}

% *** ALIGNMENT PACKAGES ***
%
\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
\ifCLASSOPTIONcompsoc
  \usepackage[caption=false,font=footnotesize,labelfont=sf,textfont=sf]{subfig}
\else
  \usepackage[caption=false,font=footnotesize]{subfig}
\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a sans serif font rather
% than the serif font used in traditional IEEE formatting and thus the need
% to invoke different subfig.sty package options depending on whether
% compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Do not use the stfloats baselinefloat ability as IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
\usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/dblfloatfix/

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley, Jeff Goldberg and 
% Axel Sommerfeldt. This package may be useful when used in conjunction with 
% IEEEtran.cls'  captionsoff option. Some IEEE journals/societies require that
% submissions have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.4.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/url/
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

\usepackage{alltt}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\usepackage{mathabx}

% for highlighting
\usepackage{color,soul}
\soulregister\cite7
\soulregister\ref7
\soulregister\pageref7

@@

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

@@

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation with Real-Time Maude}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc 
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

\hide{
\author{Jiaxiang~Liu,
        Min~Zhou,
        Xiaoyu~Song,
        Ming~Gu,
        and~Jiaguang~Sun% <-this % stops a space
\IEEEcompsocitemizethanks{%
\IEEEcompsocthanksitem J. Liu is with the School of Software, Tsinghua
University, Beijing 100084, China, and also with LIX, \'Ecole
Polytechnique, Palaiseau 91120, France.\protect\\ E-mail:
jiaxiang.liu@hotmail.com
\IEEEcompsocthanksitem M. Zhou, M. Gu and J. Sun are with the School
of Software, Tsinghua University, Beijing 100084, China.
\IEEEcompsocthanksitem X. Song is with the Department of Electrical \&
Computer Engineering, Portland State University, Portland, OR
97207-0751, USA.}% <-this % stops an unwanted space
}
%\thanks{Manuscript received April 19, 2005; revised September 17, 2014.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.
}

% The paper headers
\markboth{IEEE Transaction on xxx,~Vol.~$x$, No.~$x$, xx~$x$}%
{Liu \MakeLowercase{\textit{et al.}}: Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation with Real-Time Maude}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% The publisher's ID mark at the bottom of the page is less important with
% Computer Society journal papers as those publications place the marks
% outside of the main text columns and, therefore, unlike regular IEEE
% journals, the available text space is not reduced by their presence.
% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2014 IEEE}
% or like this to get the Computer Society new two part style.
%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2014 IEEE}%
%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer Society\hfill}}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark (Computer Society jorunal
% papers don't need this extra clearance.)

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
\begin{abstract}
Rate-Monotonic Scheduling (RMS) is one of the most important real-time
scheduling used in industry. There are a large number of results about
RMS, especially on its schedulability. However, the theoretical
results do not contain enough details to be used directly for an
industrial RMS implementation. On the other hand, the correctness of
such an implementation is of the crucial importance. In this paper, we
analyze a realistic RMS implementation by using Real-Time Maude, a
formal modeling language and analysis tool based on rewriting
logic. Overhead and some details of the hardware are taken into
account in the model. We validate the schedulability and the
correctness of the implementation within key scenarios. The soundness
and the completeness of our approach are substantiated.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
  Real-time systems, scheduling, embedded systems, modeling, formal
  verification, rewriting logic.
\end{IEEEkeywords}
}

% make the title area
\maketitle

% To allow for easy dual compilation without having to reenter the
% abstract/keywords data, the \IEEEtitleabstractindextext text will
% not be used in maketitle, but will appear (i.e., to be "transported")
% here as \IEEEdisplaynontitleabstractindextext when the compsoc 
% or transmag modes are not selected <OR> if conference mode is selected 
% - because all conference papers position the abstract like regular
% papers do.
\IEEEdisplaynontitleabstractindextext
% \IEEEdisplaynontitleabstractindextext has no effect when using
% compsoc or transmag under a non-conference mode.

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

%\IEEEraisesectionheading{\section{Introduction}\label{s:introduction}}
\section{Introduction}\label{s:introduction}
% Computer Society journal (but not conference!) papers do something unusual
% with the very first section heading (almost always called "Introduction").
% They place it ABOVE the main text! IEEEtran.cls does not automatically do
% this for you, but you can achieve this effect with the provided
% \IEEEraisesectionheading{} command. Note the need to keep any \label that
% is to refer to the section immediately after \section in the above as
% \IEEEraisesectionheading puts \section within a raised box.

% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps (small caps for compsoc).
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{P}{eriodic} task scheduling is one of the most important
topics within the field of industrial real-time systems.
%due to the large number of control systems that require cyclic activities~\cite{buttazzo2011hard}. 
A set of periodic tasks is said to be \emph{schedulable} with respect
to some scheduling algorithm if all jobs meet their
deadlines. \emph{Rate-Monotonic Scheduling} (\emph{RMS}) is a
\emph{fixed} priority scheduling algorithm for preemptive hard
real-time environments proposed by Liu and
Layland~\cite{DBLP:journals/jacm/LiuL73}, which assigns priorities to
jobs according to the periods of the corresponding tasks: the smaller
period, the higher priority. RMS is proven to be the \emph{optimal}
fixed priority scheduling algorithm~\cite{DBLP:journals/jacm/LiuL73},
in the sense that any set of tasks, which is schedulable under
\emph{some} fixed priority scheduling algorithm, is also schedulable
with respect to RMS. It is widely used in safety-critical real-time
applications, such as vehicles and avionics, thanks to its optimality
and easiness to implement.

@@

\subsection{Rate-Monotonic Scheduling Algorithm}
\label{ss:rms}
%\subsection{Formulation of the Standard Setting}
A task set consists of \emph{only} $n$ periodic tasks
$\tau_1,\ldots,\tau_n$. Each task $\tau_i$ has a period $T_i$ and a
computation requirement $C_i$. First jobs of all tasks are assumed to
be initiated at time $0$ simultaneously.  Deadlines consist of
runnability constraints only: the deadline of a job corresponding to
$\tau_i$ is the initiation time of the next job corresponding to
$\tau_i$.  The RMS algorithm chooses the labeling such that $T_1\le
T_2\le \ldots \le T_n$. Consequently, $\tau_i$ receives priority $i$,
assuming smaller numbers have higher priorities. The following
assumptions are made:

@@

\subsubsection{Specification}
Real-Time Maude models systems as \emph{modules}. A module specifies a
\emph{real-time rewrite theory} ${\cal R} = (\Sigma, E\cup A ,
\mathit{IR}, \mathit{TR})$, where:
\begin{itemize}
\item $\Sigma$ is an algebraic \emph{signature}, that is, a set of
  declarations of \emph{sorts}, \emph{subsorts} and \emph{function
    symbols}. The function symbols are allowed to be mixfix, in which
  case underscores ``\verb|_|'' indicate the positions of parameters.
  \emph{Terms} are expressions built from function symbols and
  variables.
\item $(\Sigma, E\cup A)$ is a \emph{membership equational logic
  theory}, with $E$ a set of \emph{conditional equations} and
  \emph{memberships} on $\Sigma$, and $A$ a set of equational axioms
  such as associativity, commutativity and identity.  $(\Sigma, E\cup
  A)$ models the system's ``static'' states as terms of some sort, and
  is equipped with a built-in specification of a sort \verb|Time|.
\item $\mathit{IR}$ is a set of \emph{labeled conditional rewrite
  rules} specifying the system's local transitions. Each rule has the
  form $[l]~:~t\rightarrow t'\mbox{ \textbf{if}
  }\bigwedge^n_{j=1}cond_j$, where each $cond_j$ is an equality
  $u_j=v_j$, and $l$ is a \emph{label}, \hl{$t,t',u_j,v_j$ are terms}. Such
  a rule specifies an \emph{instantaneous transition}, without
  consuming time, from an instance of $t$ to the corresponding
  instance of $t'$, \emph{provided} the conditions hold.
\item $\mathit{TR}$ is a set of (\emph{labeled}) \emph{tick rules} of
  the form $[l]~:~\{s\}\rightarrow\{s'\} \mbox{\textbf{ in time}
  }r\mbox{ \textbf{if} }cond$ that specify \emph{timed
    transitions}. \hl{Each tick rule advances time by $r$ time units from
  the \emph{entire} state modeled by term $s$ to the destination state
  $s'$.}
\end{itemize}

\hl{$\mathit{IR}$ and $\mathit{TR}$ together model the ``dynamic''
behaviors of the system.}

@@

In this paper, we only consider Real-Time Maude's \emph{Linear
  Temporal Logic (LTL) model checker}, which analyzes whether
\emph{each} behavior satisfies a temporal logic formula. \emph{State
  propositions} are defined as terms of sort \verb|Prop|. Their
semantics is defined by conditional equations of the form $\texttt{ceq
} statePattern \texttt{ |= } prop \texttt{ = } b \texttt{ if } cond$,
with $b$ a term of sort \verb|Bool|, stating that $prop$ evaluates to
$b$ in states which are instances of $statePattern$ provided the
condition $cond$ holds. These equations together define $prop$ to hold
in all states $s$ that make $s \texttt{ |= } prop$ evaluate to
\verb|true|. A temporal logic \emph{formula} is constructed by state
propositions and \hl{temporal logic operators such as} \verb|~|\hl{(negation),}
\verb|\/|\hl{(disjunction),} \verb|[]|\hl{(``always''),}
\verb|U|\hl{(``until'')}. Real-Time Maude supports both \emph{timed} and
\emph{untimed LTL model checking}. The untimed model checking command
\begin{alltt}
  (mc \(s\) |=u \(\Phi\) .)
\end{alltt}
checks whether the temporal logic formula $\Phi$ holds in all
behaviors starting from the initial state $s$, \emph{with no time
  limit}.

@@

\hide{
\begin{figure*}[!t]
\centering
\begin{picture}(300,120)(0,-10)
\thicklines
\put(0,0){\vector(1,0){320}}
\put(305,5){time(${\mu}s$)}
\thinlines
\put(0,0){\line(0,1){110}}
\put(-25,97){task$0$}
\put(-25,77){task$1$}
\put(-25,57){task$2$}
\put(-50,37){scheduling}
\put(-45,17){switching}
\put(-2,-8){$0$}

\dashline[30]{3}(0,100)(250,100)
\dashline[30]{3}(0,80)(280,80)
\dashline[30]{3}(0,60)(300,60)
\dashline[30]{3}(0,40)(210,40)
\dashline[30]{3}(0,20)(285,20)

\dashline[30]{3}(10,-10)(10,100)
\thicklines
\textcolor{red}{\put(0,40){\line(1,0){10}}}
\put(8,-18){$9$}

\thinlines
\dashline[30]{3}(50,0)(50,100)
\thicklines
\textcolor{blue}{\put(10,100){\line(1,0){40}}}
\put(35,-8){$2509$}

\thinlines
\dashline[30]{3}(55,-10)(55,80)
\thicklines
\textcolor{red}{\put(50,20){\line(1,0){5}}}
\put(46,-18){$2513$}

>>>>

%\documentclass[onecolumn]{IEEEtranTIE}
\documentclass[journal]{IEEEtranTIE}
\usepackage{graphicx}
\usepackage[noadjust]{cite}
\usepackage{picinpar}
\usepackage{amsmath}
\usepackage{stfloats}
%%\usepackage{dblfloatfix}
\usepackage{url}
\usepackage{flushend}
\usepackage[latin1]{inputenc}
\usepackage{colortbl}
\usepackage{soul}
\usepackage{multirow}
\usepackage{pifont}
\usepackage{color}
\usepackage{alltt}
\usepackage[hidelinks]{hyperref}
\usepackage{enumerate}
\usepackage{siunitx}
\usepackage{breakurl}
\usepackage{epstopdf}
\usepackage{pbox}

%% My packages
%% \usepackage[cmex10]{amsmath}
%% \interdisplaylinepenalty=2500
\usepackage{algpseudocode}
%% \usepackage{array}
\usepackage[caption=false,font=footnotesize]{subfig}

\usepackage{tikz}
\usetikzlibrary{arrows.meta}
%% \usepackage{mathabx}

@@

@@

\begin{document}
\title{Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation with Real-Time Maude}

\author{
  \vskip 1em
  Jiaxiang~Liu,
  Min~Zhou,
  Xiaoyu~Song, \emph{Member}, \emph{IEEE},
  Ming~Gu,
  and Jiaguang~Sun

  \thanks{
    Manuscript received November 8, 2015; revised March 30, 2016 and
    July 13, 2016; accepted October 20, 2016. This work was supported
    in part by NSFC Program (No. 91218302, No. 61527812), National
    Science and Technology Major Project (No. 2016ZX01038101), MIIT
    IT funds (Research and application of TCN key technologies) of
    China, and The National Key Technology R\&D Program
    (No. 2015BAG14B01-02).

    J. Liu is with the Department of Computer Science \& Technology,
    Tsinghua University, Beijing 100084, China, and also with LIX,
    \'Ecole Polytechnique, Palaiseau 91120, France (e-mail:
    jiaxiang.liu@hotmail.com).

    M. Zhou is with the School of Software, Tsinghua University,
    Beijing 100084, China (corresponding author; phone:
    +86-xxx-xxxx-xxxx; fax: xxx-xxxx-xxx; e-mail: zhoumin@gmail.com).

    X. Song is with the Department of Electrical \& Computer
    Engineering, Portland State University, Portland, OR 97207-0751,
    USA.

    M. Gu and J. Sun are with the School of Software, Tsinghua
    University, Beijing 100084, China.
  }
}

\maketitle

\begin{abstract}
Rate-monotonic scheduling (RMS) is one of the most important real-time
scheduling used in industry. There are a large number of results about
RMS, especially on its schedulability. However, the theoretical
results do not contain enough details to be used directly for an
industrial RMS implementation. On the other hand, the correctness of
such an implementation is of the crucial importance. In this paper, we
analyze a realistic RMS implementation by using Real-Time Maude, a
formal modeling language and analysis tool based on rewriting
logic. Overhead and some details of the hardware are taken into
account in the model. We validate the schedulability and the
correctness of the implementation within key scenarios. The soundness
and the completeness of our approach are substantiated.
\end{abstract}

\begin{IEEEkeywords}
Embedded systems, formal verification, modeling, real-time systems,
rewriting logic, scheduling.
\end{IEEEkeywords}

\markboth{IEEE TRANSACTIONS ON INDUSTRIAL ELECTRONICS}%
{}

\definecolor{limegreen}{rgb}{0.2, 0.8, 0.2}
\definecolor{forestgreen}{rgb}{0.13, 0.55, 0.13}
\definecolor{greenhtml}{rgb}{0.0, 0.5, 0.0}

\section{Introduction}\label{s:introduction}

\IEEEPARstart{P}{eriodic} task scheduling is one of the most important
topics within the field of industrial real-time systems. A set of
periodic tasks is said to be \emph{schedulable} with respect to some
scheduling algorithm if all jobs meet their
deadlines. \emph{Rate-monotonic scheduling} (\emph{RMS}) is a
\emph{fixed} priority scheduling algorithm for preemptive hard
real-time environments proposed by Liu and
Layland~\cite{DBLP:journals/jacm/LiuL73}, which assigns priorities to
jobs according to the periods of the corresponding tasks: the smaller
period, the higher priority. RMS is proven to be the \emph{optimal}
fixed priority scheduling algorithm~\cite{DBLP:journals/jacm/LiuL73},
in the sense that any set of tasks, which is schedulable under
\emph{some} fixed priority scheduling algorithm, is also schedulable
with respect to RMS. It is widely used in safety-critical real-time
applications, such as vehicles and avionics, thanks to its optimality
and easiness to implement.

@@

\subsection{Rate-Monotonic Scheduling Algorithm}
\label{ss:rms}
A task set consists of \emph{only} $n$ periodic tasks
$\tau_1,\ldots,\tau_n$. Each task $\tau_i$ has a period $T_i$ and a
computation requirement $C_i$. First jobs of all tasks are assumed to
be initiated at time $0$ simultaneously.  Deadlines consist of
runnability constraints only: the deadline of a job corresponding to
$\tau_i$ is the initiation time of the next job corresponding to
$\tau_i$.  The RMS algorithm chooses the labeling such that $T_1\le
T_2\le \ldots \le T_n$. Consequently, $\tau_i$ receives priority $i$,
assuming smaller numbers have higher priorities. The following
assumptions are made:

@@

\subsubsection{Specification}
Real-Time Maude models systems as \emph{modules}. A module specifies a
\emph{real-time rewrite theory} ${\cal R} = (\Sigma, E\cup A ,
\mathit{IR}, \mathit{TR})$, where:
\begin{itemize}
\item $\Sigma$ is an algebraic \emph{signature}, that is, a set of
  declarations of \emph{sorts}, \emph{subsorts} and \emph{function
    symbols}. The function symbols are allowed to be mixfix, in which
  case underscores ``\verb|_|'' indicate the positions of parameters.
  \emph{Terms} are expressions built from function symbols and
  variables.
\item $(\Sigma, E\cup A)$ is a \emph{membership equational logic
  theory}, with $E$ a set of \emph{conditional equations} and
  \emph{memberships} on $\Sigma$, and $A$ a set of equational axioms
  such as associativity, commutativity and identity.  $(\Sigma, E\cup
  A)$ models the system's ``static'' states as terms of some sort, and
  is equipped with a built-in specification of a sort \verb|Time|.
\item $\mathit{IR}$ is a set of \emph{labeled conditional rewrite
  rules} specifying the system's local transitions. Each rule has the
  form $[l]~:~t\rightarrow t'\mbox{ \textbf{if}
  }\bigwedge^n_{j=1}cond_j$, where each $cond_j$ is an equality
  $u_j=v_j$, and $l$ is a \emph{label}, $t,t',u_j,v_j$ are terms. Such
  a rule specifies an \emph{instantaneous transition}, without
  consuming time, from an instance of $t$ to the corresponding
  instance of $t'$, \emph{provided} the conditions hold.
\item $\mathit{TR}$ is a set of (\emph{labeled}) \emph{tick rules} of
  the form $[l]~:~\{s\}\rightarrow\{s'\} \mbox{\textbf{ in time}
  }r\mbox{ \textbf{if} }cond$ that specify \emph{timed
    transitions}. Each tick rule advances time by $r$ time units from
  the \emph{entire} state modeled by term $s$ to the destination state
  $s'$.
\end{itemize}

$\mathit{IR}$ and $\mathit{TR}$ together model the ``dynamic''
behaviors of the system.

@@

In this paper, we only consider Real-Time Maude's \emph{Linear
  Temporal Logic (LTL) model checker}, which analyzes whether
\emph{each} behavior satisfies a temporal logic formula. \emph{State
  propositions} are defined as terms of sort \verb|Prop|. Their
semantics is defined by conditional equations of the form $\texttt{ceq
} statePattern \texttt{ |= } prop \texttt{ = } b \texttt{ if } cond$,
with $b$ a term of sort \verb|Bool|, stating that $prop$ evaluates to
$b$ in states which are instances of $statePattern$ provided the
condition $cond$ holds. These equations together define $prop$ to hold
in all states $s$ that make $s \texttt{ |= } prop$ evaluate to
\verb|true|. A temporal logic \emph{formula} is constructed by state
propositions and temporal logic operators such as \verb|~|(negation),
\verb|\/|(disjunction), \verb|[]|(``always''),
\verb|U|(``until''). Real-Time Maude supports both \emph{timed} and
\emph{untimed LTL model checking}. The untimed model checking command
\begin{alltt}
  (mc \(s\) |=u \(\Phi\) .)
\end{alltt}
checks whether the temporal logic formula $\Phi$ holds in all
behaviors starting from the initial state $s$, \emph{with no time
  limit}.

@@

\subsection{Evaluation}
We now show in this section that our results are both sound and
complete.
